---
title: "Movie ratings"
output: html_document
date: "2026-02-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R ny embedded R code chunks within the document. You can embed an R code chunk like this:Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of a




Introduction

This project analyzes a small movie dataset to identify genre-wise rating trends and overall rating distribution using R. The analysis includes aggregation, sorting, and data visualization using dplyr and ggplot2.

Load Libraries

```{r}
library(dplyr)
library(ggplot2)
```

Create dataset
```{r}
movies <- data.frame(
  Title = c("The Truman Show", "Inception", "The Notebook", "Titanic",
            "Pride and Prejudice", "John Wick", "Interstellar",
            "La La Land", "Harry Potter"),
  
  Genre = c("Drama", "Sci-Fi", "Romance", "Romance",
            "Romance", "Action", "Sci-Fi",
            "Romance", "Fantasy"),
  
  Rating = c(8.2, 9.0, 7.5, 8.0,
             8.1, 7.8, 9.2,
             8.0, 8.5)
)

head(movies)
```

```{r}
str(movies)
summary(movies)
```

```{r}
genre_summary <- movies %>%
  group_by(Genre) %>%
  summarise(
    Average_Rating = mean(Rating),
    Movie_Count = n()
  ) %>%
  arrange(desc(Average_Rating))

genre_summary
```
Visualization: Average Rating by Genre
```{r}
ggplot(genre_summary, aes(x = reorder(Genre, Average_Rating), 
                          y = Average_Rating, 
                          fill = Genre)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Average Rating by Genre",
       x = "Genre",
       y = "Average Rating") +
  theme_minimal() +
  theme(legend.position = "none")
```
Identify Top Rated Movie
```{r}
top_movie <- movies %>%
  arrange(desc(Rating)) %>%
  slice(1)

top_movie
```
Rating Distribution
```{r}
ggplot(movies, aes(x = Rating)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Movie Ratings",
       x = "Rating",
       y = "Count") +
  theme_minimal()
```

Conclusion

The analysis shows that Sci-Fi has the highest average rating among the selected movies. Interstellar is the top-rated movie. Most movie ratings fall between 7.5 and 9.5, indicating generally strong audience reception across genres.